I want to express my gratitude to my advisor, Prof. Harjinder Singh, for helping me take my first steps research, help explore the many domains I might be interested in and eventually for helping me start making progress on Quantum Computing. I would also like to thank my senior Utkarsh Azad for mentoring me through the entire process, from formulating the problem to benchmarking the results, on both the projects and for procing the requisite domain knowledge and intuition required to tackle those problems. Both Prof. Harjinder Singh and Utkarsh Azad have also greatly helped with writing and editing this thesis.

Next I would like to thank the professors in the CCNSB and CQST departments for the design of courses that helped me gain interest in the sciences, and specifically in Quantum. Prof. Subhadip Mitra, Prof. Deva Priyakumar, Prof. Santanan Chakraborty, and many others have taken courses which have helped me hone my skills in this domain.

I would also like to thank my friends helped a lot with techniques learnt and ideas developed in the project. Bhuvanesh Sridharan helped out a lot with the Monte Carlo Tree Search formulation of the problem and helped find ways to iteratively improve on the project. Kalp Shah and Jai Bardhan contributed valuable thoughts on both scientific and machine learning aspects of the projects involved. Other friends and lab-mates including Adrian Alva, Shweta Sahoo, Gaurang Tandon, Kanish Anand, and many others have engaged in discussion on these topics which have been invaluable to me in validating and extending my ideas.

Finally, I would like to thank my Mom and Dad for their constant support and guidance on all matters including academic and technical advice. 
