This dissertation shows that Quantum-inspired Machine-Learning methods can be naturally applied in various parts of the Quantum processing pipeline. We present two novel methods: qRoute - a reinforcement-learning-based quantum circuit routing method, and qLEET - a variational quantum circuit visualization and evaluation framework. These methods and results are one small step on the way of improving the capability of near-term quantum computers to a point useful quantum-simulations can be executed. Hamiltonian simulations for Quantum chemistry is likely one of the tasks that is possible to execute even on noisy computers, and our experiments and results have been motivated along these lines.

In qRoute, we have shown state-of-the-art performance on various quantum circuit compilation tasks. This demonstrates that reinforcement learning settings can learn algorithms more performant than manually engineered heuristics. The routing task is one step of the long hardware-efficient algorithm development and compilation pipeline. The solution for the routing task is also a proposed solution to the general problem of decision over subsets.

In qLEET, we have shown a framework consisting of both optimization-inspired and quantum-inspired methods to evaluate the learning capabilities of variational quantum circuits. These visualized metrics include loss landscape, training trajectories, expressibility, entanglement capability, and parameter histogram of these circuits. The framework presented is provided with code compatible with Cirq, Qiskit, and PyTket quantum circuits and works with any choice of the loss function, optimizer, and noise model. Our method helps guide ansatz design and parametrization better for solving these variational problems.
The software package for qRoute has been made publically available on \url{https://github.com/AnimeshSinha1309/qroute-router/}.

The routing process is just one part of the compilation process, where we convert a circuit that does not respect the constraints of the hardware to one that does. Converting a unitary directly to a hardware-constraint-respecting circuit and optimizing noise under a specific noise model instead of minimizing depth are goals that we need to design algorithms for, and qRoute seems like a viable solution candidate as a part of the pipeline for these processes. There is also scope for adaptation of the qRoute algorithm for these objectives. Similarly, qLEET can be extended with more visualizations and analyses of the optimization landscape and quantum resource utilization. qLEET has been explored on a small class of algorithms like QAOA and VQE, visualization of other algorithms and variation algorithm design and hyperparameter tuning using qLEET has been left as future scope of the project. 
The software package for qLEET has been made publically available on \url{https://github.com/QLemma/qleet}.

In the entirety, we believe that the contributions of this dissertation will help better the time-efficiency and noise-immunity of some algorithms on noisy near-term intermediate-scale quantum computers and will serve as some inspiration to pipelines for other tasks in near-term quantum algorithm design.
